---
title: "Data Basics & Sampling Principles <br><br> MTH 361 A/B <br> Applied Statistics I <br> Spring 2023"
author: University of Portland
date: "2022-01-23"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# R Packages {.bigger}

```{r message=FALSE, warning=FALSE, paged.print=FALSE, echo=TRUE}
library(tidyverse)
library(openintro)
```

R packages are groups of R functions, compiled code, and example data. In the R environment, they are saved in the `library` directory. R installs a set of packages by default after installation. More packages will be added later if they are required for a specific purpose. When we launch the R console, just the default packages are available. Other packages that are already installed must be explicitly loaded before they may be utilized by the R application.

Throughout this presentation, we are using the [`county`](https://www.openintro.org/data/index.php?data=county){target="_blank"} dataset.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# remove NAs in teh county dataframe
county <- county %>% na.omit()
```

# Loading Datasets {.bigger}

```{r loading-data-from-package, message=FALSE, echo=TRUE}
# selecting only 3 rows to display from the county dataframe
county %>%
  slice(1:3)
```

Examples of variable types.

1. `name` - nominal categorical variable
2. `state` - nominal categorical variable
3. `pop2000` - discrete numerical variable
4. `pop_change` - continuous numerical variable
5. `median_edu` - ordinal categorical variable

# Accessing Rows and Columns {.bigger}

  * Using brackets to access **specific indices** of the data frame with the first 3 rows.
```{r rows-and-columns-1, message=FALSE, echo=TRUE}
# accessing only the 1st 3 rows and 1st 5 columns of the county dataframe
county[1:3,1:5]
```

  * Using brackets to access **specific column labels** of the data frame with the first 3 rows.
```{r rows-and-columns-2, message=FALSE, echo=TRUE}
# accessing only the 1st 3 rows and four specific variables 
#   from the county dataframe
county[1:3,c("name","state","pop2000","pop2010")]
```

# Accessing Entire Columns {.bigger}

  * Using the `$` command to access a **specific column label** of the data.
```{r rows-and-columns-3, message=FALSE, echo=TRUE}
# accessing all rows of the pop2000 variable from the county dataframe
glimpse(county$pop2000)
```

  * Using brackets to access **entire columns** using **specific column labels**.
```{r rows-and-columns-4, message=FALSE, echo=TRUE}
# accessing all rows of three specific variables  from the county dataframe
glimpse(county[,c("pop2000","pop2010","pop2017")])
```

# Basic Descriptive Statistics {.bigger}

  * Using `$` command to **summarize one column**.
```{r summarize-data-1, message=FALSE, echo=TRUE}
# computing basic statistics of the pop2000 variable from the county dataframe
summary(county$pop2000)
```

  * Using brackets to **summarize multiple columns**.
```{r summarize-data-2, message=FALSE, echo=TRUE}
# computing basic statistics of three specific columns from the county dataframe
summary(county[,c("pop2000","pop2010","pop2017")])
```

  * Computing **Mean and Standard Deviation of one column**.
```{r means-and-standard-deviations-1, message=FALSE, echo=TRUE}
# computing the mean and standard deviation of the pop2000 variable 
#   from the county dataframe
mean(county$pop2010)
sd(county$pop2010)
```

# Histograms {.bigger}

```{r basic-data-visualization-1, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
# plot a histogram of the pop2000 variable 
#   from the county dataframe
ggplot(data = county, aes(x = pop2000)) + geom_histogram(bins=60)
```

# Scatterplots {.bigger}

```{r basic-data-visualization-2, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
# plot a scatter plot of the variables unemployment_rate and poverty
#   from the the county dataframe
ggplot(data = county, aes(x = unemployment_rate, y = poverty)) + geom_point()
```

# Simple Random Sampling {.bigger}

*Example:* Suppose we want to randomly sample 100 observations (rows) from the data frame.

```{r echo=TRUE}
# number of samples
n <- 100

# pick random row indices from the data frame
#   the replace=FALSE means that we sampling without replacement
rand_index <- sample.int(dim(county)[1],n, replace = FALSE)

# select rows using the random indices
simple_random_sample <- county %>% slice(rand_index)

# take a glimpse of the random sample
glimpse(simple_random_sample)
```

# Stratified Random Sampling {.bigger}

*Example 1:* Suppose we want to sample 50\% of observations of each state in the data frame.

```{r echo=TRUE}
strat_sample_frac <- county %>%
                     # group by state (states as strata)
                     group_by(state) %>%
                     # create new column that counts 
                     #    the number of observations in each strata
                     mutate(num_rows=n()) %>% 
                     # randomly sample 50% of observations in each strata
                     sample_frac(0.50, weight=num_rows) %>% 
                     # make data frame ungrouped
                     ungroup()

# take a glimpse of the random sample
glimpse(strat_sample_frac)
```

Now, let's check if the code did it right.

```{r echo=TRUE}
# checking number of total observations
nto <- dim(county %>% filter(state == "California"))[1]
# checking number of sample observations
nso <- dim(strat_sample_frac %>% filter(state == "California"))[1]

# ratio
nso/nto # the ratio should be 0.50
```

*Example 2:* Suppose we want to sample exactly 100 observations in each level of the `metro` variable.

```{r echo=TRUE}
stratified_sample <- county %>% 
                     # group by metro (metro - yes or no - as strata)
                     group_by(metro) %>% 
                     # randomly sample 100 observations in each strata
                     sample_n(100) %>% 
                     # make data frame ungrouped
                     ungroup()

# take a glimpse of the random sample
glimpse(stratified_sample)
```

# Clustered Random Sampling {.bigger}

*Example:* Suppose we consider each state as a cluster, and we randomly sample 10 clusters.

```{r echo=TRUE}
# number of clusters to randomly sample
n_clusters <- 10

# get list of states
states <- levels(county$state)

# sample the states
states_sample <- sample(states,n_clusters)

# get the rows with corresponding states that are sampled.
cluster_sample <- county %>% filter(state==states_sample) 

# take a glimpse of the random sample
glimpse(cluster_sample)
```
